---
title: "Linear Regression"
author: "Sriram Vivek"
output: pdf_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)       
library(MASS)        
library(leaps)       
library(dplyr)       

```
Using the random seed 123 to divide the data into 75% training and 25% testing.

```{r }

ames_data <- read.csv("/Users/sriram/Desktop/SEMESTER 2/AMS 580/Linear Regression/Ames_Housing_Data.csv")


ames_data <- na.omit(ames_data)


set.seed(123)


index <- createDataPartition(ames_data$SalePrice, p = 0.75, list = FALSE)
train_data <- ames_data[index, ]
test_data <- ames_data[-index, ]

```

Finding the best model using the stepwise variable selection method (based on the
BIC criterion) using the training data and displaying necessary information.

       
```{r }


full_model <- lm(SalePrice ~ ., data = train_data)


step_model <- stepAIC(full_model, direction = "both", k = log(nrow(train_data)), trace = FALSE)


cat("Stepwise Model Coefficients:\n")
print(coef(step_model))


step_predictions <- predict(step_model, newdata = test_data)


step_rmse <- sqrt(mean((test_data$SalePrice - step_predictions)^2))
step_r_squared <- cor(test_data$SalePrice, step_predictions)^2

cat("\nStepwise Model Performance:\n")
cat("RMSE:", step_rmse, "\n")
cat("R^2:", step_r_squared, "\n")

```

Finding the best model using the best subset variable selection method (based on
the SSE criterion) and siplaying the necessary information.

```{r }



subset_model <- regsubsets(SalePrice ~ ., data = train_data, nvmax = 20)

best_subset_index <- which.min(summary(subset_model)$bic)
best_subset_vars <- names(coef(subset_model, id = best_subset_index))[-1]  # Exclude intercept

best_subset_formula <- as.formula(paste("SalePrice ~", paste(best_subset_vars, collapse = " + ")))
best_subset_model <- lm(best_subset_formula, data = train_data)

cat("\nBest Subset Model Coefficients:\n")
print(coef(best_subset_model))

best_subset_predictions <- predict(best_subset_model, newdata = test_data)

best_subset_rmse <- sqrt(mean((test_data$SalePrice - best_subset_predictions)^2))
best_subset_r_squared <- cor(test_data$SalePrice, best_subset_predictions)^2

cat("\nBest Subset Model Performance:\n")
cat("RMSE:", best_subset_rmse, "\n")
cat("R^2:", best_subset_r_squared, "\n")

```

Comparison of the above two models
           
```{r }


step_bic <- BIC(step_model)
best_subset_bic <- BIC(best_subset_model)

cat("\nModel Comparison (BIC):\n")
cat("Stepwise Model BIC:", step_bic, "\n")
cat("Best Subset Model BIC:", best_subset_bic, "\n")

cat("\nModel Comparison (Test Data):\n")
cat("Stepwise Model RMSE:", step_rmse, "\n")
cat("Best Subset Model RMSE:", best_subset_rmse, "\n")
cat("Stepwise Model R^2:", step_r_squared, "\n")
cat("Best Subset Model R^2:", best_subset_r_squared, "\n")

```






